GG_topi

### 2018-11-26 cage_effect.Rmd
26112018

---
title: "Cage effect"
author: "Francesco Beghini"
date: "November 8, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

>>>>

---
title: "Cage effect"
author: "Francesco Beghini"
date: "November 8, 2018"
output: 
  html_document: 
    df_print: kable
    theme: cerulean
editor_options: 
  chunk_output_type: console
---

### 2018-11-26 da_test.Rmd
26112018

---
title: "Differential Analysis"
output: 
  html_document: 
    number_sections: yes
    theme: cosmo
    toc: yes
editor_options: 
  chunk_output_type: console
---

@@

### Boxplot of OTUs abundances of differenitally abundant OTUs
```{r}
  t37_samples_idx <- grep('T7|T3', sample_names(ggtopi_filt), perl = TRUE)
  tax_table(gg_topi.relab)[rownames(res37.filt),]

@@

  genus_ab_t %>% bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=interaction(timepoint, condition))) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme_bw() +
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

>>>>

---
title: "Differential Analysis"
output: 
  html_document: 
    theme: cosmo
    toc: yes
editor_options: 
  chunk_output_type: console
---

@@

@@

  genus_ab_t %>%
    bind_rows(genus_ab_t[genus_ab_t$timepoint=='T7',-5]) %>%
    bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
    ggplot() +
      geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=interaction(timepoint, condition))) + 
      facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
      coord_flip() + 
      theme_bw() +
      theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
             strip.text.y = element_text(angle = 0))
```

### 2018-11-26 gg_topi.Rmd
26112018

---
title: "GG_topi"
author: "Francesco Beghini"
date: "26/09/2018"
output: 
  html_document: 
    df_print: kable
    self_contained: no
    theme: cerulean
editor_options: 
  chunk_output_type: inline
---

@@

distancesjac <- reshape2::melt(as.matrix(distjac), varnames = c('rows','cols'))
distancesjac$condizione1 <- sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][2])
distancesjac$condizione2 <- sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][2])
distancesjac$topo1 <- factor(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][3]))
distancesjac$topo2 <- factor(sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][3]))
distancesjac$timepoint1 <- stringr::str_replace(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distancesjac$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
```

@@

ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
  geom_point(size = 0.8) +
  geom_point(data=centroids, size = 4)
```

@@

# Distribution of distances between conditions per timepoint
```{r}
dct <- distances %>%
  filter(timepoint1 == timepoint2) %>%
  select(timepoint1, condizione1, condizione2, value) %>%
  dplyr::rename(timepoint = timepoint1) %>% 
  group_by(timepoint, condizione1, condizione2) %>%
  summarise(mediandistance=median(value)) 
```
```{r include=TRUE,  results = "asis"}
for(t in unique(dct$timepoint)){
  print(dct %>%
    filter(timepoint == t) %>% 
    dplyr::rename(condizione = condizione2) %>%
    spread(condizione1, mediandistance) %>%
    knitr::kable())

@@

>>>>

---
title: "GG_topi"
author: "Francesco Beghini"
date: "26/09/2018"
output: 
  html_document: 
    df_print: kable
    theme: cosmo
editor_options: 
  chunk_output_type: console
---

@@

distancesjac <- reshape2::melt(as.matrix(distjac), varnames = c('rows','cols'))
distancesjac$condizione1 <- sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][2])
distancesjac$condizione2 <- sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][2])
distancesjac$topo1 <- factor(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][3]))
distancesjac$topo2 <- factor(sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][3]))
distancesjac$timepoint1 <- stringr::str_replace(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distancesjac$timepoint2 <- stringr::str_replace(sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
```

@@

ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
  geom_point(size = 0.8) +
  geom_point(data=centroids, size = 4) + 
  theme_minimal()
```

@@

# Distribution of distances between conditions per timepoint
```{r}
dct <- distances %>%
  filter(timepoint1 == timepoint2) %>%
  select(timepoint1, condizione1, condizione2, value) %>%
  dplyr::rename(timepoint = timepoint1) %>% 
  group_by(timepoint, condizione1, condizione2) %>%
  summarise(mediandistance=median(value)) 
```

```{r include=TRUE,  results = "asis"}
for(t in unique(dct$timepoint)){
  print(dct %>%
    filter(timepoint == t) %>% 
    dplyr::rename(condizione = condizione2) %>%
    spread(condizione1, mediandistance) %>%
    knitr::kable())

@@

# Abundance of Bacteroidales 24/4 through timepoints
```{r}
idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")

otu_table(gg_topi.genus.relab)[idx_247,] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(timepoint, Abundance, fill=condition)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    # coord_flip() + 
    theme_bw() +
    theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
           strip.text.y = element_text(angle = 0))
```

# Correlation of tumor size at T7 and abundance of Bacteroidales 24/ through timepoints
```{r}
mass_247 <- otu_table(gg_topi.genus.relab)[idx_247,] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  mutate(mouseID=as.integer(stringr::str_sub(.$SampleID,8, 10))) %>%
  left_join(mappingfile[mappingfile$timepoint=='T7',], by=c('mouseID' = 'id-topo'))

mass_247 %>%
  ggplot(aes(Abundance,tumor_mass)) +
  geom_point() + 
  facet_grid(timepoint.x~condition, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
ds_corr_247 <- data.frame()
for(x in unique(mass_247$timepoint.x)){
  for(y in unique(mass_247$condition)){
    tmp_res <- with(mass_247 %>% filter(timepoint.x==x & condition==y), cor.test(Abundance, tumor_mass))
    ds_corr_247 %<>% bind_rows(data.frame(Cond = paste(y, ' ', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
  }
}
print(knitr::kable(ds_corr_247))
```

# Correlation of tumor size at T7 and difference of abundance of Bacteroidales 24/7 through timepoints
```{r}
mass_247 %>%
  filter(timepoint.x %in% c('T3','T7')) %>%
  select(Abundance, timepoint.x, condition, tumor_mass,mouseID) %>%
  tidyr::spread(timepoint.x, Abundance) %>%
  mutate(T3_7 = abs(T7-T3)) %>%
  ggplot(aes(T3_7,tumor_mass)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
with(mass_247 %>%
  filter(timepoint.x %in% c('T3','T7')) %>%
  select(Abundance, timepoint.x, condition, tumor_mass,mouseID) %>%
  tidyr::spread(timepoint.x, Abundance) %>%
  mutate(T3_7 = abs(T7-T3)),
  cor.test(T3_7,tumor_mass))

## Ordination: T3 vs T7 with tumor size
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T7','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(shape=timepoint, color=tumor_mass), size=3, alpha=.7, stroke=0.8) +
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .1, ymax = .1)
```
